# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
library(timeDate)

rawdata<-read.csv("activity.csv")
data1<-rawdata[!is.na(rawdata$steps),]
hist1data<-aggregate( steps ~ date , data = data1 , sum )

mean1data<-mean(hist1data$steps)
median1data<-median(hist1data$steps)
```

## What is mean total number of steps taken per day?

Mean total number of steps taken per day is **`r mean1data` steps** and median is **`r median1data` steps**.

## Histogram of data for given time without missing values

```{r fig.height=4, fig.width=11}
barplot(hist1data$steps,names.arg=hist1data$date, cex.names=0.6, main="Histogram of steps per day (NA removed) ")
```
## What is the average daily activity pattern?
```{r fig.height=4, fig.width=11}
time1data<-aggregate( steps ~ interval , data = data1 , mean )
maxsteps<-max(time1data$steps)
maxindex<-time1data[which.max(time1data$steps),1]
plot(time1data, type="l", main="Daily activity pattern (NA removed)")
```
Maximum average number of steps is **`r maxsteps` steps** and it is in interval **"`r maxindex`" interval**.

## Imputing missing values
Missing data will be replaced by average number of steps for particular interval based on previous calculation.

```{r}
newdata<-rawdata
missingdata<-sum(is.na(rawdata$steps))
i<-0
for (x in newdata$steps){
        i<-i+1
        if(is.na(x)){
                temp<-i%%nrow(time1data)
                if (temp == 0){
                        newdata[i,1]<-time1data[1,2]        
                        }
                else {
                        newdata[i,1]<-time1data[temp,2]
                        }
                }
        }
```
There is all together **`r missingdata`** missing values

```{r}
hist2data<-aggregate( steps ~ date , data = newdata , sum )
mean2data<-mean(hist2data$steps)
median2data<-median(hist2data$steps)
```
```{r fig.height=4, fig.width=11}
barplot(hist2data$steps,names.arg=hist2data$date, cex.names=0.6, main="Histogram of steps per day (NA replaced with average) ")
```
Mean total number of steps taken per day with filled NA according to given formula is **`r mean2data` steps** and median is **`r median2data` steps**.

Difference is **`r mean2data-mean1data` steps** in favour of "replaced NA data" and difference in median is **`r median2data-median1data` steps** in favour of "replaced NA data".

## Are there differences in activity patterns between weekdays and weekends?

### Preprocessing the data
```{r}
timevector<-newdata$date
timevector<- strptime(timevector, "%Y-%m-%d", tz = "")
newtimevector<-c()
for (i in 1:length(timevector)){
        if(isWeekday(timevector[i], wday = 1:5)){
                newtimevector[i]<-"weekday"
                }
        else {
                newtimevector[i]<-"weekend"
                }
        }
newtimevector<-as.factor(newtimevector)
newdata<-cbind(newdata, newtimevector)

```
### Extracting the data
```{r fig.height=4, fig.width=11}
weekdays <- subset(newdata, newtimevector=="weekday", select=c(steps, interval))
plotweekdays<-aggregate( steps ~ interval , data = weekdays , mean )

weekend <- subset(newdata, newtimevector=="weekend", select=c(steps, interval))
plotweekend<-aggregate( steps ~ interval , data = weekend , mean )

plotdifference<-plotweekend
plotdifference$steps<-plotdifference$steps-plotweekdays$steps

sumweekdays<-sum(plotweekdays$steps)
sumweekends<-sum(plotweekend$steps)

```
### Panel plot
```{r fig.height=11, fig.width=11}
par(mfrow=c(3,1), mar=c(1,4,3,1))
plot(plotweekdays, type="l", main="Weekdays",ylim=c(0, 250))
plot(plotweekend, type="l", main="Weekends", ylim=c(0, 250))
plot(plotdifference, type="l", main="Difference weekends-weekdays", ylim=c(-125, 125))
```
Average number of steps in one day durong weekdays is **`r sumweekdays` steps** and on weekend is **`r sumweekends` steps**.

From this we can conclude that **activity is higher during weekend** and comparing the graphs, **activity resulting in diffenerce is mainly occuring during afternoon hours**. **During weekdays the peak of activity is 8 in the morning**, and generaly **during weekdays is higher activity in early morning.**
